---
title: "The Issue With Single Issuer Districts"
author: "Sam Powers"
date: "9/3/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message=FALSE, widgetframe_self_contained = FALSE, widgetframe_isolate_widgets = TRUE, widgetframe_widgets_dir = 'widgets')

# library(sf)
library(plotly)
library(ggthemes)
library(urbnmapr)
library(ggiraph)
library(Cairo)
source("https://gist.githubusercontent.com/benmarwick/2a1bb0133ff568cbe28d/raw/fb53bd97121f7f9ce947837ef1a4c65a73bffb3f/geom_flat_violin.R")
library(knitr)
library(networkD3)
library(treemap)
library(d3treeR)
library(widgetframe)
library(htmlwidgets)
library(tidyverse)


# link between rating areas and counties

state_regions <- read_csv("State_Regions.csv")

county_rating <- read_csv("individual_county_rating_area_crosswalk_2019_2019_04_01.csv") %>%
  mutate(county_fips = paste0("0",fips_code)) %>%
  separate(county_fips, c(NA, "county_fips"), -5)

issuers_counties   <- read_csv("individual_issuer_county_report_2019_2019_08_29.csv") %>%
  mutate(county_fips = paste0("0",fips_code)) %>%
  separate(county_fips, c(NA, "county_fips"), -5)


number_issuers_counties <- issuers_counties %>%
  filter(market == "on_market") %>%
  group_by(state, county_name, county_fips) %>%
  summarize(count_issuers = n(), count_plans = sum(plan_count)) %>% ungroup()

us_counties_graph <- urbnmapr::counties %>% select(-county_name) %>% inner_join(number_issuers_counties, by = "county_fips") %>%
  select(county_name, county_fips, everything())

us_states_regions <-  state_regions %>% inner_join( urbnmapr::states , by = c("ST"= "state_abbv"))

 
 
issuers_rating_area <-
number_issuers_counties %>%
  inner_join(county_rating) %>%
  select(AREA = rating_area_id, count_issuers, count_plans, county_name) %>%
  mutate(SINGLEISSUER = ifelse(count_issuers == 1, 1, 0))


plans_regular <- read_csv("plan_data.csv")

nrow(plans_regular)




issuers_by_area  <-
plans_regular %>%
  select(ST, CARRIER, AREA) %>%
  inner_join(state_regions) %>%
  distinct() %>%
  group_by(Region, ST, AREA) %>%
  summarize(ISSUERS = n(), SINGLEISSUER = ifelse(ISSUERS == 1, 1, 0 ))

single_issuer_descriptors <-
plans_regular %>%
  inner_join(issuers_by_area) %>%
    filter( METAL == "Silver" , actively_marketed == TRUE) %>%
  select(SINGLEISSUER, everything())
 


```


Have some intro here about the relevance of this topic within healthcare. Maybe contextualize it within the current political discussion, idk... more research needed. 


```{r number_issuers, fig.width = 20, fig.height = 15}

plans_per_county <- 
us_counties_graph %>%
  mutate(tooltip = gsub( "'","", paste0( county_name, ", ", state_name, ": <br> ", count_plans, " plans"))) %>%

  ggplot(aes(long, lat, group = group, fill = count_plans)) +
  geom_polygon_interactive(color = "white", size = .1, aes( tooltip = tooltip)) +
  scale_fill_gradientn(colors = c("white", "dodgerblue2", "blue4"), 
                     limits = c(-10,  300),
                     breaks = c(0,50,100, 150,200, 250, 300),
                      guide = guide_colorbar(
                        #labesl = scales::dollar,
                         direction = "horizontal",
                                            title.position = "top", nbin = 10)
                      ) +
  geom_polygon(data = states, mapping = aes(long, lat, group = group ),
              fill = NA, color = "#ffffff") +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  theme(legend.title = element_text(),
        legend.key.width = unit(.5, "in")
        ) +
  labs(fill = "")+    
  theme(panel.grid.major = element_blank(), 
        panel.background = element_rect(fill = "white"),
        legend.key.width = unit(2.5, "line"),
        legend.key.height = unit(1, "line"),
        legend.position = "top",
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
       plot.title = element_text(size=22, hjust = .5 )

        ) +
  ggtitle("Number of Plans Operating Per County")

ggiraph(code = {print(plans_per_county)},  hover_css = "fill:red;")

```



At first blush, it looks like there are plenty of healthcare plans to go around. Some states, like New York, have way more plans than others. But even each county in Nebraska has nine plans to choose from. 

```{r single_issuers, fig.width = 20, fig.height = 15}
single_issuers <-
us_counties_graph %>%
  mutate(single = ifelse(count_issuers == 1, 1, 0),
         tooltip = gsub( "'","", paste0( county_name, ", ", state_name ))) %>%

ggplot(aes(long, lat, group = group, fill = single)) +
  geom_polygon_interactive( color = "white", size = .1, aes( tooltip = tooltip)) +
  scale_fill_gradientn(colors = c("white", "maroon"), 
                      limits = c(-.1,  1),
                   #   breaks = c(1,2,3, 4, 5, 6, 7, 7, 8, 9),
                      guide = guide_colorbar(
                        #labesl = scales::dollar,
                         direction = "horizontal",
                                            title.position = "top", nbin = 10)
                      ) +
  geom_polygon(data = states, mapping = aes(long, lat, group = group),
              fill = NA, color = "#ffffff") +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  theme(legend.title = element_text(),
        legend.key.width = unit(.5, "in")
        ) +
  labs(fill = "Number of Issuers on Individual Market in Rating Area")+    
  theme(panel.grid.major = element_blank(), 
        panel.background = element_rect(fill = "white"),
        legend.key.width = unit(2.5, "line"),
        legend.key.height = unit(1, "line"),
        legend.position = "none",
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        plot.title = element_text(size=22, hjust = .5 )

        ) +
  ggtitle("Districts With Only One Issuer")

ggiraph(code = {print(single_issuers)},  hover_css = "fill:red;")
```

But then we start to break it down. And we find that there are so many districts where one issuer (Insurance Company) provides all of the care for the entire population. The Northeast reports no single issuer states. The issue primarily affects the South and parts of the Midwest and West. 



# Minimum Price for a Silver Plan in the Rating Area
```{r traits, fig.width = 15, fig.height = 10}
single_issuer_descriptors %>%
  inner_join(county_rating, by  = c("AREA" = "rating_area_id")) %>% 
    mutate(SINGLE = ifelse(SINGLEISSUER == 1, "Single Issuer for County", "Multiple Issuers in County")) %>%
  group_by(Region, SINGLE, ST, AREA, county_name) %>%
  summarize(minimum = min(PREMI27) ) %>%
  ungroup() %>%
  group_by(Region, SINGLE) %>%
  summarize(minimum = round(mean(minimum)) ) %>%
   mutate( minimum = paste0("$",minimum, "/Month") )%>%
spread(SINGLE, minimum) %>%
  kable()


```

One main issue with single-issuer districts is that it reduces competition and variability in care options. The "Silver" Plan is considered the average plan that every American should have. It covers all the basic domains of healthcare without being exhorbitantly priced. And yet, we find, that in single-issuer districts, the cheapest silver plan available can be, on average, over $100 more per month than the cheapest silver plan in a multi-issuer district in the Midwest, about $90 more in the South, and more than $200 more in the West 


The distributions themselves of minimum available price points for Silver Plans are very different. 

``` {r,  fig.width = 8, fig.height = 6}

raincloud_theme = theme(
  text = element_text(size = 10),
  axis.title.x = element_text(size = 16),
  axis.title.y = element_text(size = 16),
  axis.text = element_text(size = 14),
  axis.text.x = element_text(angle = 45, vjust = 0.5),
  legend.title=element_text(size=16),
  legend.text=element_text(size=16),
  legend.position = "right",
  plot.title = element_text(lineheight=.8, face="bold", size = 22, hjust = .5, vjust = 1),
  panel.border = element_blank(),
  panel.grid.minor = element_blank(),
  panel.grid.major = element_blank(),
  axis.line.x = element_line(colour = 'black', size=0.5, linetype='solid'),
  axis.line.y = element_line(colour = 'black', size=0.5, linetype='solid'))

single_prices <-
single_issuer_descriptors %>%
  inner_join(county_rating, by  = c("AREA" = "rating_area_id")) %>% 
    mutate(SINGLE = ifelse(SINGLEISSUER == 1, "Single Issuer", "Multiple Issuers")) %>%

 filter(Region != "Northeast" ) %>%
  group_by(Region, SINGLE, ST, county_name) %>%
  summarize(mimimum = min(PREMI27) ) %>%
  
  ggplot( aes(x = SINGLE, y = mimimum, fill = SINGLE )) +
  geom_flat_violin(position = position_nudge(x = .2, y = 0), alpha = .8)+
  geom_point(aes(y = mimimum, color = SINGLE), position = position_jitter(width = .15), size = .5, alpha = 0.8) +
  geom_boxplot(width = .1,  outlier.shape = NA, alpha = 0.5) +
  expand_limits(x = 1) +
  guides(fill = FALSE) +
  guides(color = FALSE) +
  # scale_color_brewer(palette = "Spectral") +
  # scale_fill_brewer(palette = "Spectral") +
  coord_flip() +
  theme_bw() +
  raincloud_theme +
  facet_grid(Region ~ .)  +

ggtitle('Minimum Price of Silver Plan in County') +

  labs(x="", y="$/Month" , fill = "")  
 
single_prices

```


In fact, we find, that when we compare within states, plans offered without Issuer competition are, on average, about $70 more expensive per month.



```{r, results = "hide"}

# This chunk's Results shouldnt actually render #

 summary(lm(PREMI27 ~ SINGLEISSUER +  ST, single_issuer_descriptors  ))

 issuer_fit <- lm(PREMI27 ~ SINGLEISSUER + ST , single_issuer_descriptors)
 state_fit <- lm(PREMI27 ~ ST , single_issuer_descriptors)

 anova(issuer_fit, state_fit, test = "LRT")
```


Digging alittle bit deeper into consumer agency, we find that Single Issuer Districts often do not provide Silver Plan subscribers the ability to choose how they would like to pay for their coverage. In fact, all 85 Single Issuer Counties in the West require Silver Plan subscribers to pay all emergency medical costs up to the price of their deductible whereas 202 of the ones in the Midwest require the same. 


```{r, fig.width = 10, fig.height = 7}

comparison_data <-
single_issuer_descriptors %>%
 inner_join(county_rating, by  = c("AREA" = "rating_area_id")) %>% 
  group_by(Region, AREA, county_name, SINGLEISSUER) %>%
  mutate(denominator = n()) %>%
  ungroup() %>%
mutate(SINGLE = ifelse(SINGLEISSUER == 1, "Single Issuer", "Multiple Issuers"),

    ER_Indicator = ifelse(ER_CopayInnTier1 %in% c(0,99), "Coinsurance","Copay"),
    ER_Application =  ifelse(ER_CopayInnTier1 %in% c(0,99) , ER_CoinsInnTier1, ER_CopayInnTier1),
    ER_Value =  ifelse(ER_CopayInnTier1 %in% c(0,99), paste0(ER_CoinsInnTier1A, "% of Bill") , paste0("$",ER_CopayInnTier1A)),
    
    ER_Description = case_when(
   ER_Application ==  1  ~ "No Charge",
  ER_Application  ==   2 ~ "Pay Everything Until Deductible",
   ER_Application ==  3 ~ "Always Pay Benefit",
   ER_Application == 4 ~ "Use Only After Paying Deductible"
    ),
  
  ER_Deductible = case_when(
     ER_Application ==  1  ~ "Not Responsible",
  ER_Application  ==   2 ~ "Full Deductible",
   ER_Application ==  3 ~ "Just Benefit",
   ER_Application == 4 ~ "Full Deductible"
   
  )
  
  ) %>%
    
  group_by(  Region, AREA, county_name, SINGLE, denominator, ER_Deductible) %>%
  summarize(count = n()) %>%
  filter(ER_Deductible == "Full Deductible") %>%
  ungroup() %>%
  mutate(percent = count/denominator *100,
         option = ifelse(percent == 100, "no", "yes")) %>%
  group_by(Region, SINGLE) %>%
  mutate(denominator = n()) %>%
  group_by(Region, denominator, SINGLE, option) %>%
  summarize(count = n()) %>%
  filter(option == "no") %>%
  mutate(percent = count/denominator*100)
  



 p <- comparison_data   %>% 
  filter(Region != "Northeast") %>%

  ggplot( aes(x = reorder(Region, percent), y = percent, fill = reorder(as.factor(SINGLE), percent))) + 
  geom_bar(stat = "identity", position = "dodge", alpha = .8, color = "black") +
  
  ggtitle('Counties that Require Silver Plan Holders to Pay Out-Of-Pocket Until Deductible for Emergency Care') +
  theme(
   plot.title = element_text( face="bold", hjust = 0.5 ),
    legend.title = element_text(),
    axis.title.x = element_text(vjust=-2),
    axis.title.y = element_text(vjust= 2),
    axis.text.x=element_text(vjust=-4.5),
    axis.ticks.x = element_blank(),
   panel.background = element_blank()

  )+ 
  scale_y_continuous(labels = function(x) paste0(round(x), "%"), expand = c(0, 0 ), limits = c(-.5, 105)) +
  geom_text(aes(label= paste0(round(percent), "%"), y = percent + 2), position = position_dodge(width = .9), size = 5  ) +
  geom_text(aes(label= paste0( "(n=", count,")"), y = 0), position = position_dodge(width = .9), vjust = 2 ) +
  labs(x="Region", y="Counties Requiring Complete Out-Of-Pocket Payment Until Deductible, %", fill = "Number of Issuers")  +
   coord_cartesian(clip = 'off') +

   scale_fill_brewer(palette="Set1")



p
```

Below, Single Issuer Counties appear in blue while counties that require out-of-pocket payment up to the price of the deductible are shaded in. You can see that almost all the Single Issuer (blue) counties are shaded, showing that, where there is only one health insurance company, consumers often are stuck without the option to choose different ways to pay for care.  

```{r, fig.width= 10, fig.height = 5}
 plot_data <-
single_issuer_descriptors %>%
 inner_join(county_rating, by  = c("AREA" = "rating_area_id")) %>% 
  group_by(Region, AREA, county_name, SINGLEISSUER) %>%
  mutate(denominator = n()) %>%
  ungroup() %>%
mutate(SINGLE = ifelse(SINGLEISSUER == 1, "Single Issuer", "Multiple Issuers"),

    ER_Indicator = ifelse(ER_CopayInnTier1 %in% c(0,99), "Coinsurance","Copay"),
    ER_Application =  ifelse(ER_CopayInnTier1 %in% c(0,99) , ER_CoinsInnTier1, ER_CopayInnTier1),
    ER_Value =  ifelse(ER_CopayInnTier1 %in% c(0,99), paste0(ER_CoinsInnTier1A, "% of Bill") , paste0("$",ER_CopayInnTier1A)),
    
    ER_Description = case_when(
   ER_Application ==  1  ~ "No Charge",
  ER_Application  ==   2 ~ "Pay Everything Until Deductible",
   ER_Application ==  3 ~ "Always Pay Benefit",
   ER_Application == 4 ~ "Use Only After Paying Deductible"
    ),
  
  ER_Deductible = case_when(
     ER_Application ==  1  ~ "Not Responsible",
  ER_Application  ==   2 ~ "Full Deductible",
   ER_Application ==  3 ~ "Just Benefit",
   ER_Application == 4 ~ "Full Deductible"
   
  )
  
  ) %>%
    
  group_by(  Region, AREA, county_name, SINGLE, denominator, ER_Deductible) %>%
  summarize(count = n()) %>%
  filter(ER_Deductible == "Full Deductible") %>%
  ungroup() %>%
  mutate(percent = count/denominator *100,
         option = ifelse(percent == 100, "Yes", "No")) %>%
  inner_join(us_counties_graph) 
  
  ggplot(  plot_data ,aes(long, lat, group = group) )+
  geom_polygon(color = "white", size = .1, aes(fill = SINGLE, alpha = as.factor(option))) +
    
   scale_color_manual(values=c( "#800000", "#2980B9"), guide = "none") +
  scale_alpha_discrete(range= c( .35,1 )) +

  geom_polygon(data = states, mapping = aes(long, lat, group = group ),
              fill = NA, color = "#ffffff") +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  theme(legend.title = element_text(),
        legend.key.width = unit(.5, "in")
        ) +
  labs(fill = "Single or Multi Issuer", alpha = "Requiring Complete Deductible Payment" )+    
  theme(panel.grid.major = element_blank(), 
        panel.background = element_rect(fill = "white"),
        legend.key.width = unit(2.5, "line"),
        legend.key.height = unit(1, "line"),
        legend.position = "right",
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
       plot.title = element_text(size=14, hjust = .5 )

        ) +
  ggtitle("Single and Multi Issuer Counties \n Requiring Complete Payment Until Deductible for All Silver Plans") +
       scale_fill_brewer(palette="Set1")



```
This is an issue because, a deductible in these single issuer counties is on average between $3300 and $4400 for an individual, depending on region, assuming the emergency center they are taken to is within their network, and assuming the doctor at that hospital is also within their network. In some of these counties, the mean deductible is as high as $6000. So imagine that, just because of where you live, you must purchase a healthcare plan that will not contribute to your emergency care  until you have spent at least $4000 out of pocket. 

### Minimum, Mean, and Maximum Average Deductibles required in Single-Issuer Counties where complete paymen is required for Emergency Care
```{r}
pay_full <-
single_issuer_descriptors %>%
 inner_join(county_rating, by  = c("AREA" = "rating_area_id")) %>% 
  group_by(Region, AREA, county_name, SINGLEISSUER) %>%
  mutate(denominator = n()) %>%
  ungroup() %>%
mutate(SINGLE = ifelse(SINGLEISSUER == 1, "Single Issuer", "Multiple Issuers"),

    ER_Indicator = ifelse(ER_CopayInnTier1 %in% c(0,99), "Coinsurance","Copay"),
    ER_Application =  ifelse(ER_CopayInnTier1 %in% c(0,99) , ER_CoinsInnTier1, ER_CopayInnTier1),
    ER_Value =  ifelse(ER_CopayInnTier1 %in% c(0,99), paste0(ER_CoinsInnTier1A, "% of Bill") , paste0("$",ER_CopayInnTier1A)),
    
    ER_Description = case_when(
   ER_Application ==  1  ~ "No Charge",
  ER_Application  ==   2 ~ "Pay Everything Until Deductible",
   ER_Application ==  3 ~ "Always Pay Benefit",
   ER_Application == 4 ~ "Use Only After Paying Deductible"
    ),
  
  ER_Deductible = case_when(
     ER_Application ==  1  ~ "Not Responsible",
  ER_Application  ==   2 ~ "Full Deductible",
   ER_Application ==  3 ~ "Just Benefit",
   ER_Application == 4 ~ "Full Deductible"
   
  )
  
  ) %>%
    
  group_by(  Region, AREA, county_name, SINGLE, denominator, ER_Deductible) %>%
  summarize(count = n(), Deductible = median(Med_Deductible_Amount, na.rm = TRUE)) %>%
  filter(ER_Deductible == "Full Deductible") %>%
  ungroup() %>%
  mutate(percent = count/denominator *100,
         pay_full = ifelse(percent == 100, "Yes", "No")) %>%
  filter(pay_full == "Yes", SINGLE == "Single Issuer") %>%
  group_by(Region) %>%
  summarize(Min = min(round(Deductible)), Median  = round( median(Deductible)), Max = round(max(Deductible)), Counties = n()) %>%
  kable()

pay_full


```




```{r, fig.width = 10, fig.height = 7}
comparison_data_2 <-
single_issuer_descriptors %>%
   inner_join(county_rating, by  = c("AREA" = "rating_area_id")) %>% 
  group_by(Region, SINGLEISSUER, AREA, county_name) %>%
  mutate(denominator = n()) %>%
  ungroup() %>%

mutate(SINGLE = ifelse(SINGLEISSUER == 1, "Single Issuer", "Multiple Issuers"),

    OM_Indicator = ifelse(OM_CopayInnTier1 %in% c(0,99), "Coinsurance","Copay"),
    OM_Application =  ifelse(OM_CopayInnTier1 %in% c(0,99) , OM_CoinsInnTier1, OM_CopayInnTier1),
    OM_Value =  ifelse(OM_CopayInnTier1 %in% c(0,99), paste0(OM_CoinsInnTier1A, "% of Bill") , paste0("$",OM_CopayInnTier1A)),
    
    OM_Description = case_when(
   OM_Application ==  1  ~ "No Charge",
  OM_Application  ==   2 ~ "Pay Everything Until Deductible",
   OM_Application ==  3 ~ "Always Pay Benefit",
   OM_Application == 4 ~ "Use Only After Paying Deductible"
    ),
  
  OM_Deductible = case_when(
     OM_Application ==  1  ~ "Not Responsible",
  OM_Application  ==   2 ~ "Full Deductible",
   OM_Application ==  3 ~ "Just Benefit",
   OM_Application == 4 ~ "Full Deductible"
   
  )
  
  ) %>%
    
  group_by(  Region, SINGLE, denominator, AREA, county_name, OM_Deductible) %>%
  summarize(count = n()) %>%
  filter(OM_Deductible == "Full Deductible") %>%
  mutate(percent = count/denominator *100,
         option = ifelse(percent == 100, "no", "yes")) %>%
  group_by(Region, SINGLE) %>%
  mutate(denominator = n()) %>%
  group_by(Region, denominator, SINGLE, option) %>%
  summarize(count = n()) %>%
  filter(option == "no") %>%
  mutate(percent = count/denominator*100) %>%
  bind_rows(
    data.frame(Region = "West", denominator = 0, SINGLE = "Multiple Issuers", option = "no", count = 0, percent = 0)
    
    
  )



 p <- comparison_data_2   %>% 
  filter(Region != "Northeast") %>%

  ggplot( aes(x = reorder(Region, percent), y = percent, fill = reorder(as.factor(SINGLE), percent))) + 
  geom_bar(stat = "identity", position = "dodge", alpha = .8, color = "black") +
  
  ggtitle(' Percent of Silver Plans Requiring Out-Of-Pocket Payment \n Until Deductible for Outpatient Mental Health Care') +
  theme(
   plot.title = element_text( face="bold", hjust = 0.5 ),
    legend.title = element_text(),
    axis.title.x = element_text(vjust=-2),
    axis.title.y = element_text(vjust= 2),
    axis.text.x=element_text(vjust=-4.5),
    axis.ticks.x = element_blank(),
   panel.background = element_blank()

  )+ 
  scale_y_continuous(labels = function(x) paste0(round(x), "%"), expand = c(0, 0 ), limits = c(-.5, 70)) +
  geom_text(aes(label= paste0(round(percent), "%"), y = percent + 1), position = position_dodge(width = .9), size = 5  ) +
  geom_text(aes(label= paste0( "(n=", count,")"), y = 0), position = position_dodge(width = .9), vjust = 2 ) +
  labs(x="Region", y="Silver Plans Requiring Complete Out-Of-Pocket Payment Until Deductible, %", fill = "Number of Issuers")  +
   coord_cartesian(clip = 'off') +

   scale_fill_brewer(palette="Set1")

p
```

<p> some text </p>
  


